\section{Related Work}

\subsection{Information Fusion, Missing Views, and Semantic Alignment}

Information fusion has long treated source identity, representation alignment, missing views, uncertainty, and late or intermediate fusion rules as separate but interacting design choices \cite{castanedo2013datafusion,lahat2015multimodal,baltrusaitis2019multimodal,zhao2017multiview}. Quantum readout gives these choices a particularly sharp form. A client readout is only a partial view of a public observable dictionary, and a coordinate is a physical measurement choice--for example a Pauli string, qubit subset, or measurement basis--rather than an arbitrary feature index. Two numerical readout vectors are therefore not safely fusible until the protocol states which physical observable each coordinate estimates.

Incomplete multi-view and missing-modality methods usually repair or bypass missing views through imputation, modality dropout, prompting, reconstruction, graph or matrix completion, or representation sharing \cite{havaei2016hemis,tang2024incomplete,cai2024projected,wu2024missingmodality}. HeMIS-style fusion, for example, learns modality-specific encoders and aggregates available-view embeddings, making it a strong full-dimensional trainable reference \cite{havaei2016hemis}. Multimodal FL adds privacy and client heterogeneity; recent surveys and benchmarks emphasize missing modalities, modality imbalance, and corrupted labels as robustness challenges \cite{lin2023federatedmultimodal,che2023multimodalsurvey,feng2023fedmultimodal,yu2024fedinmm}. We borrow these baselines deliberately. Schema zero-fill corresponds to simple imputation, mask-only asks whether missingness metadata is enough, shared-observable aggregation keeps only the intersection of views, and HeMIS checks the case where a trainable missing-modality architecture sees full feature-level inputs. Quantum readout differs because missingness is induced by measurement design, not by an absent entry in a table. Observable identity must therefore be part of the protocol itself.

This distinction is also why the comparison is not framed as a claim that QProto replaces learned missing-modality models. Many such models are powerful once a centralized training set provides feature-level samples whose modality identities are already known. QProto targets an earlier protocol question: before reconstruction, modality dropout, or late fusion is meaningful, do the transmitted coordinates have shared source identities, and can compact client statistics be fused without confusing a missing coordinate with a measured zero? This makes schema/order binding and coverage counts first-class fusion objects rather than preprocessing details.

\subsection{Federated Learning as Distributed Fusion}

Federated learning is commonly introduced as decentralized optimization, but it is also a privacy-constrained information-fusion problem: the server combines partial client evidence without direct access to the underlying data sources \cite{yang2019federated,kairouz2021advances}. Classical FL usually assumes that the quantities being combined have compatible semantics, for example shared parameters, shared labels, calibrated sensors, or aligned feature spaces. Heterogeneous FL weakens this assumption through non-IID labels, different models, missing modalities, feature anchors, and client-specific feature spaces \cite{li2020fedprox,karimireddy2020scaffold,acar2021feddyn,reddi2021fedopt,rakotomamonjy2024flic,zhou2024fedfa}. In heterogeneous QFL, the units being fused are finite-shot estimates of quantum observables, so the semantic identity of each coordinate is a physical measurement choice. \method\ makes this identity explicit through a public readout schema.

\subsection{Federated Optimization and Prototype Aggregation}

FedAvg remains the standard communication-efficient reference point for FL \cite{mcmahan2017fedavg}. FedProx stabilizes local training with a proximal term \cite{li2020fedprox}; SCAFFOLD uses control variates to reduce client drift \cite{karimireddy2020scaffold}; FedDyn aligns local and global objectives through dynamic regularization \cite{acar2021feddyn}; and FedOpt covers adaptive server optimizers such as FedAdam, FedYogi, and FedAdagrad \cite{reddi2021fedopt}. These methods are about optimization drift or non-IID data. They do not by themselves solve the more basic question of whether the exchanged coordinates refer to the same source.

Prototype-based FL aggregates class representations rather than gradients or full parameters. FedProto shows that class prototypes can handle heterogeneous client models \cite{tan2022fedproto}. Related representation-level FL methods rely on shared embedding dimensions, public heads, or distillation signals. Our setting is stricter: even if two clients upload same-dimensional class prototypes, coordinate \(a\) may correspond to different quantum observables. Thus \method\ should be viewed as a semantic comparability layer for prototype fusion in QFL, not merely as another prototype averaging rule.

\subsection{Quantum Federated Learning}

Quantum FL now includes hybrid quantum-classical models, QNNs, sequence models, quantum data, and privacy-preserving workflows \cite{chen2021fqml,xia2021quantumfed,chehimi2022qfl,chehimi2024fedqlstm,innan2024fedqnn,kumar2023privacy,ren2025towardqfl,ballester2025qflreview}. Most of this literature asks how to train quantum models or aggregate quantum model parameters. We include trainable PennyLane VQC FedAvg/FedProx baselines for that setting, but the question here is different: what may the server aggregate when clients expose different measurement subsets, readout orders, and finite-shot statistics? Such readout-level heterogeneity can arise for VQCs, quantum kernels, shadow-style measurements, or hybrid QNNs.

\subsection{Quantum Kernels, Readout, and Measurement Sketches}

Quantum kernel methods compare data through quantum feature maps and can offer useful inductive biases when the induced geometry is hard to approximate classically \cite{havlicek2019supervised,schuld2019feature,huang2021power}. Classical shadows estimate many quantum observables from randomized measurements and have become a major tool for efficient property prediction \cite{huang2020classicalshadows,hadfield2022locallybiased,bu2024pauli}. Readout-error mitigation and shot-noise-aware estimation similarly emphasize that measurement outcomes are not generic features; they are noisy estimates tied to a specified measurement procedure. \method\ is complementary to these lines. It does not introduce a new quantum feature map, shadow protocol, or mitigation method. Instead, it specifies how heterogeneous readout statistics should be bound to public observable identities before federated fusion.

\subsection{Kernel Mean Embeddings and Random Features}

Kernel mean embeddings represent probability distributions as expectations in reproducing kernel Hilbert spaces \cite{smola2007hilbert,gretton2012mmd}. Random Fourier features approximate shift-invariant kernels with finite-dimensional sketches \cite{rahimi2007random}. \method\ uses these tools to aggregate class-conditional adapted readout distributions with bounded communication. The \hopmethod\ extension adds projected quadratic features, which are useful when second-order class structure matters. The distinction from ordinary KME aggregation is the schema adapter: without a public value-and-mask domain, the empirical distributions being embedded are not distributions over the same space.

\input{tables/table_method_positioning}

\subsection{Positioning}

Table~\ref{tab:method-positioning} summarizes the distinction in protocol terms rather than in broad method names. The closest conceptual neighbors are FedProto, feature-heterogeneous FL, missing-view imputation, mask-aware deep fusion, late fusion, quantum parameter-FL, and measurement-sketch methods. Each covers part of the design space. The novelty of \method\ is not that value-mask vectors or prototypes are individually new; it is the combination required when source identities themselves are heterogeneous: explicit schema/order binding, missing-zero separation, non-overlapping but semantically identified sources, finite-shot communication accounting, and wrong-schema controls. Without schema binding, prototype averaging may be numerically well-defined but semantically meaningless as a fusion operation.
