\section{Discussion and Limitations}

\subsection{Main Empirical Message for Information Fusion}

The main empirical lesson is that comparability has to be specified before fusion. Quantum readout makes the point concrete, but nothing in the failure mode is uniquely quantum: the same ambiguity appears when clients expose source-specific coordinates, local orders, missing views, or unequal measurement budgets. The no-schema, forced-canonical, wrong-schema, zero-fill, and mask-only controls correspond to familiar fusion failures--uncalibrated sources, incorrect alignment, simple imputation, and metadata-only fusion. Their behavior is the reason we treat the schema as part of the protocol rather than as bookkeeping.

The real-data benchmarks put that claim outside the quantum setting. UCI HAR, UCI Multiple Features, MHEALTH, PAMAP2, Hydraulic Systems, and WDBC cover phone sensors, wearable blocks, industrial condition monitoring, multi-view digit features, and biomedical measurements. In these settings, coverage-aware QProto is competitive with PCA, ridge and autoencoder imputation, mask-aware pooled MLP, and view-wise late fusion, while it is consistently stronger than equal-byte shared-view RFF controls. HeMIS is stronger on several tasks when trained as a full-dimensional missing-modality classifier; we view this as a useful boundary, not a contradiction, because QProto is designed for compact protocol statistics under source-identity and payload constraints. Put differently, QProto is not a substitute for HeMIS or MLP-style missing-modality learners when centralized full-dimensional training is available. It addresses the complementary protocol question that those models usually take for granted: whether distributed source coordinates have shared semantic identities and can be communicated as compact federated statistics. Multiple Features shows why source groups should not be treated as anonymous coordinates when view dimensionalities differ. MHEALTH is the clearest real interaction-heavy case for adaptive CHOP. PAMAP2, Hydraulic Systems, MFeat, and HAR are the counterweight: there, low-order or calibrated mixtures are safer than blind HOP. WDBC is nearly saturated. Together these cases give a narrower but cleaner claim: QProto is a schema/value/count fusion protocol for settings where source identity, coverage, and communication constraints would otherwise make aggregation ambiguous.

\subsection{Role of HOP}

The high-order module should be used selectively. The full-class image experiments and the real UCI HAR/MFeat/PAMAP2/WDBC tasks are largely low-order or coverage-separable after schema alignment; CProto or group-balanced low-order fusion is usually the safer compact statistic. MHEALTH, the PennyLane high-order readout, the covariance-sensing task, and the synthetic covariance benchmark contain stronger second-order or cross-source structure, and mask-aware HOP becomes more useful there. Adaptive CHOP turns this observation into a rule: it improves over pure CHOP on HAR, MFeat, and PAMAP2, and on the ten-seed MHEALTH run it gives the highest mean among compact equal-byte statistics by choosing non-CProto candidates on 6/10 seeds. The key is that HOP is still part of the comparability protocol. Random-key HOP is weak on UCI HAR and WDBC and gives no gain over random-key CProto on MHEALTH; on MFeat, random-key low-order CProto is itself strong, so we do not use that benchmark to claim key-selection superiority. Cross-source CHOP helps only when anchor-selected keys correspond to meaningful source interactions.

\subsection{Role of Shot-Variance Shrinkage}

Shot-variance shrinkage remains a reliability hook rather than a headline method. In highly non-IID label splits, downweighting noisy clients can also suppress rare class evidence. A more useful version would be class-aware or calibration-aware, instead of a single client-level reliability scalar.

\subsection{Hardware and Backend Limitations}

The implementation currently covers backend-agnostic simulated readouts, PennyLane circuit-generated readouts, and Qiskit Aer noisy readouts. These are compatibility checks, not hardware validation. A small IBM hardware demonstration with the same schema adapter and a 2-class or 4-class task would be a natural next step if low-cost access is available. The code already includes a Qiskit circuit/count-to-readout path: it builds one circuit per observable, converts bit counts into expectation values, and writes the same precomputed-readout format used by the main experiments. In this environment the path was verified on Aer fallback only, because IBM Runtime is not installed and no token is available.

\subsection{Dataset Scope}

The dataset coverage is broad but not exhaustive. The experiments include first-four-class diagnostics and full ten-class experiments for MNIST, Fashion-MNIST, and CIFAR-10 features, scale and observable-overlap sweeps, real UCI HAR, UCI Multiple Features, MHEALTH, PAMAP2, Hydraulic Systems, and WDBC missing-view benchmarks, and controlled non-image high-order fusion tasks. Gas Drift appears only in the supplement as a boundary check: even with cross-sensor HOP pair selection, low-order or learned baselines are stronger, so it is not used as HOP evidence. Remote sensing or a larger industrial monitoring benchmark would further reduce the chance that the paper is read as primarily quantum-computing-specific.
