\documentclass[preprint]{elsarticle}

\usepackage{amsmath,amssymb,amsfonts}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{xcolor}
\usepackage{url}
\usepackage{array}
\usepackage{amsthm}
\usepackage{etoolbox}
\usepackage{placeins}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,fit,backgrounds,calc}

\graphicspath{{figures/}}

\newcommand{\method}{QProto}
\newcommand{\hopmethod}{QProto-HOP}
\newcommand{\schema}{\mathsf{S}}
\newcommand{\adapter}{A_{\schema}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\cD}{\mathcal{D}}
\newcommand{\cO}{\mathcal{O}}
\newcommand{\cC}{\mathcal{C}}
\newcommand{\cK}{\mathcal{K}}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\setlength{\tabcolsep}{4pt}
\renewcommand{\arraystretch}{1.08}
\setlength{\bibsep}{0pt}
\newcommand{\qtabfont}{\scriptsize}
\AtBeginEnvironment{tabular}{\qtabfont}
\AtBeginEnvironment{tabular*}{\qtabfont}
\sloppy

% Suppress the default elsarticle preprint footer in the review PDF.
\makeatletter
\def\ps@pprintTitle{%
  \let\@oddhead\@empty
  \let\@evenhead\@empty
  \let\@oddfoot\@empty
  \let\@evenfoot\@empty}
\makeatother

\journal{Information Fusion}

\begin{document}

\begin{frontmatter}

\title{Protocol-Defined Semantic Fusion for Heterogeneous and Incomplete Source Coordinates}

\author[inst1]{Leyuan Zhang}
\ead{leyuan.zhang@durham.ac.uk}
\author[inst1]{Yang Long\corref{cor1}}
\ead{yang.long@durham.ac.uk}
\cortext[cor1]{Corresponding author.}
\address[inst1]{Department of Computer Science, Durham University, Durham, United Kingdom}

\begin{abstract}
Distributed fusion often treats coordinates as if their meanings were already shared. In heterogeneous and incomplete source settings this assumption can fail: clients may expose different coordinates, view subsets, local orders, noise levels, or measurement budgets. Federated quantum readout is a concrete stress case because a coordinate is a measurement choice, but the same issue appears in sensor and multi-view fusion. We study this as a semantic comparability problem rather than another averaging rule. \method\ binds local coordinates to a public schema, maps observations into a value-mask domain, and aggregates class-wise statistics with coverage counts. The default low-order backbone separates missing measurements from measured near-zero values and normalizes source groups with unequal dimensionality. For covariance- or interaction-dominated tasks, \hopmethod/CHOP adds compressed mask-aware cross-source outer-product statistics, with a training-only gate that falls back to CProto when high-order evidence is not useful. We evaluate quantum-readout image tasks, real missing-view benchmarks, Qiskit Aer/PennyLane readouts, and covariance-sensing tasks under matched schemas, payloads, or feature access. CProto reaches 0.753 on UCI HAR, versus 0.344/0.391 for equal-byte zero-fill/shared-view RFF. Group-balanced CProto reaches 0.876 on Multiple Features. Adaptive CHOP reaches 0.878 on subject-held-out MHEALTH, and compact CHOP reaches 0.977 on Hydraulic Systems. Full-dimensional HeMIS or MLP models can be stronger, and HOP helps mainly when source interactions matter. Overall, source-coordinate fusion requires named coordinates, not anonymous positions.
\end{abstract}

\begin{keyword}
Source-coordinate fusion \sep semantic comparability \sep incomplete views \sep schema-bound prototypes \sep heterogeneous federated learning \sep high-order sketches
\end{keyword}

\end{frontmatter}

\input{sections/01_introduction}
\input{sections/02_related_work}
\input{sections/03_problem}
\input{sections/04_method}
\input{sections/05_analysis}
\input{sections/06_experiments}
\input{sections/07_discussion}
\input{sections/08_conclusion}

\FloatBarrier
\bibliographystyle{elsarticle-num}
\bibliography{references}

\end{document}


